Main Analysis

### 2024-05-27 Manuscript Analysis/adaptation_analysis.Rmd
added model comparison

summary(rep_acc_mod)

@@

sjPlot::tab_model(m5_ple, m6_ple, m7_ple, m8_ple,
                  pred.labels = c("Intercept", "Trial Order", "Speaker [ALS]", "Repetition Accuracy [Inaccurate]",
                                  "Speaker [ALS] * Repetition Accuracy [Inaccurate]", "Trial Order * Speaker [ALS]",
                                  "Trial Order * Repetition Accuracy [Inaccurate]", "Trial Order * Speaker * Repetition Accuracy"),
                  dv.labels = c("Model 5", "Model 6", "Model 7", "Model 8"),
                 file = "Tables/ple_adaptation_models.html")

@@

m0_pupil <- glmmTMB(peak_pupil ~ 1 + (1|subject), data = max_pupil)
summary(m0_pupil)
performance::icc(m0_pupil)

@@

m1_pupil <- glmmTMB(peak_pupil ~ 1 + (trial_c|subject), data = max_pupil)
summary(m1_pupil)
performance::test_performance(m0_pupil, m1_pupil)

@@

m2_pupil <- glmmTMB(peak_pupil ~ trial_c + (trial_c|subject), data = max_pupil)
summary(m2_pupil)
performance::test_performance(m1_pupil, m2_pupil)

@@

m3_pupil <- glmmTMB(peak_pupil ~ trial_c + speaker + (trial_c|subject), data = max_pupil)
summary(m3_pupil)
performance::test_performance(m2_pupil, m3_pupil)

@@

m4_pupil <- glmmTMB(peak_pupil ~ trial_c + speaker + rep_acc + (trial_c|subject), data = max_pupil)
summary(m4_pupil)
performance::test_performance(m3_pupil, m4_pupil)

@@

m5_pupil <- glmmTMB(peak_pupil ~ trial_c + speaker + rep_acc + speaker*rep_acc + (trial_c|subject), data = max_pupil)
summary(m5_pupil)
performance::test_performance(m4_pupil, m5_pupil)

@@

m5b_pupil <- glmmTMB(peak_pupil ~ speaker + rep_acc + speaker*rep_acc + (trial_c|subject), data = max_pupil)
summary(m5b_pupil)
performance::test_performance(m5_pupil, m5b_pupil)

@@

m6_pupil <- glmmTMB(peak_pupil ~ trial_c + speaker + rep_acc + 
                      speaker*rep_acc + trial_c*speaker +(trial_c|subject), data = max_pupil)
summary(m6_pupil)
performance::test_performance(m4_pupil, m6_pupil)

@@

m7_pupil <- glmmTMB(peak_pupil ~ trial_c + speaker + rep_acc + 
                      speaker*rep_acc + trial_c*speaker + trial_c*rep_acc + (trial_c|subject), data = max_pupil)
summary(m7_pupil)
performance::test_performance(m6_pupil, m7_pupil)

@@

m8_pupil <- glmmTMB(peak_pupil ~ trial_c + speaker + rep_acc + 
                      speaker*rep_acc + trial_c*speaker + trial_c*rep_acc + 
                      trial_c*speaker*rep_acc + (trial_c|subject), data = max_pupil)
summary(m8_pupil)
performance::test_performance(m7_pupil, m8_pupil)

@@

adapt_pupil <- bam(normed_pupil ~ condition +
                  s(time_norm, by = condition, k = 25) +
                  te(time_norm, trial_c, k = 25) +
                  # Random Smooth for listener
                  s(time_norm, listener, bs = 'fs', m =1),
                discrete = T,
                family = "scat",
                AR.start = start_event,
                rho = .92,
                data = pupil_df)

summary(adapt_pupil)

@@

>>>>

@@

sjPlot::tab_model(m5_ple, m6_ple, m7_ple, m8_ple, show.ci = F,
                  pred.labels = c("Intercept", "Trial Order", "Speaker [ALS]", "Repetition Accuracy [Inaccurate]",
                                  "Speaker [ALS] * Repetition Accuracy [Inaccurate]", "Trial Order * Speaker [ALS]",
                                  "Trial Order * Repetition Accuracy [Inaccurate]", "Trial Order * Speaker * Repetition Accuracy"),
                  dv.labels = c("Model 5", "Model 6", "Model 7", "Model 8"),
                 file = "Tables/ple_adaptation_models.html")

@@

m0_pupil <- glmmTMB(peak_pupil ~ 1 + (1|subject), data = max_pupil)
sjPlot::tab_model(m0_pupil)
performance::icc(m0_pupil)

@@

m1_pupil <- glmmTMB(peak_pupil ~ 1 + (trial_c|subject), data = max_pupil)
sjPlot::tab_model(m1_pupil)
performance::test_performance(m0_pupil, m1_pupil)

@@

m2_pupil <- glmmTMB(peak_pupil ~ trial_c + (trial_c|subject), data = max_pupil)
sjPlot::tab_model(m2_pupil)
performance::test_performance(m1_pupil, m2_pupil)

@@

m3_pupil <- glmmTMB(peak_pupil ~ trial_c + speaker + (trial_c|subject), data = max_pupil)
sjPlot::tab_model(m3_pupil)
performance::test_performance(m2_pupil, m3_pupil)

@@

m4_pupil <- glmmTMB(peak_pupil ~ trial_c + speaker + rep_acc + (trial_c|subject), data = max_pupil)
sjPlot::tab_model(m4_pupil)
performance::test_performance(m3_pupil, m4_pupil)

@@

m5_pupil <- glmmTMB(peak_pupil ~ trial_c + speaker + rep_acc + speaker*rep_acc + (trial_c|subject), data = max_pupil)
sjPlot::tab_model(m5_pupil)
performance::test_performance(m4_pupil, m5_pupil)

@@

m5b_pupil <- glmmTMB(peak_pupil ~ speaker + rep_acc + speaker*rep_acc + (trial_c|subject), data = max_pupil)
sjPlot::tab_model(m5b_pupil)
performance::test_performance(m5_pupil, m5b_pupil)

@@

m6_pupil <- glmmTMB(peak_pupil ~ trial_c + speaker + rep_acc + 
                      speaker*rep_acc + trial_c*speaker +(trial_c|subject), data = max_pupil)
sjPlot::tab_model(m6_pupil)
performance::test_performance(m4_pupil, m6_pupil)

@@

m7_pupil <- glmmTMB(peak_pupil ~ trial_c + speaker + rep_acc + 
                      speaker*rep_acc + trial_c*speaker + trial_c*rep_acc + (trial_c|subject), data = max_pupil)
sjPlot::tab_model(m7_pupil)
performance::test_performance(m6_pupil, m7_pupil)

@@

m8_pupil <- glmmTMB(peak_pupil ~ trial_c + speaker + rep_acc + 
                      speaker*rep_acc + trial_c*speaker + trial_c*rep_acc + 
                      trial_c*speaker*rep_acc + (trial_c|subject), data = max_pupil)
sjPlot::tab_model(m8_pupil)
performance::test_performance(m7_pupil, m8_pupil)

@@

adapt_pupil <- bam(normed_pupil ~ condition +
                  s(time_norm, by = condition, k = 30) +
                  te(time_norm, trial_c, k = 30) +
                  # Random Smooth for listener
                  s(time_norm, listener, bs = 'fs', m =1),
                discrete = T,
                family = "scat",
                AR.start = start_event,
                rho = .92,
                data = pupil_df)

summary.gam(adapt_pupil)

@@

vis.gam(adapt_pupil, view = c("time_norm", "trial_c"), 
        cond = list(condition = "Control Accurate"), 
        theta = -45, phi = 55, color = "topo")

vis.gam(adapt_pupil, view = c("time_norm", "trial_c"), 
        cond = list(condition = "Control Inaccurate"), 
        theta = -45, phi = 55, color = "topo")

vis.gam(adapt_pupil, view = c("time_norm", "trial_c"), 
        cond = list(condition = "ALS Accurate"), 
        theta = -45, phi = 55, color = "topo")

vis.gam(adapt_pupil, view = c("time_norm", "trial_c"), 
        cond = list(condition = "ALS Inaccurate"), 
        theta = -45, phi = 55, color = "topo")
```

### 2024-05-27 Manuscript Analysis/main_analysis.Rmd
added model comparison

library(rio)
library(tidyverse)
library(mgcv)
library(itsadug)
library(irr)
library(ggpubr) # installl.packages("ggpubr")
library(gt) # install.packages("gt")
library(gtsummary) # install.packages("gtsummary")
library(glmmTMB) # install.packages("glmmTMB")
library(performance) # install.packages("performance")
library(sjPlot) # install.packages("sjPlot")

@@

Adding in AR(1) structure to address autocorrelation concerns. Bumped up knots too.

@@

# summary(m1_pupil)

m1_prediction <- plot_smooth(m1_pupil,
            view = "time_norm",
            plot_all = "condition",
            rug = F, se = 1, rm.ranef = T, plot = F)

@@

In this model, the scaled t distribution was added to the model specifications to fix model fit to the data. The model was originally running into an error when the random smooth for phrase was included in the specification. Therefore, it was removed in this model. The random smooth for listener remained in this model

m2_pupil <- bam(normed_pupil ~ condition +
                  s(time_norm, by = condition, k = 25) +
                  # Random Smooths for listener
                  s(time_norm, listener, bs = 'fs', m =1),
                discrete = T,
                AR.start = start_event,
                rho = start_value_rho(m0_pupil),
                family = "scat",
                data = pupil_df)

# summary(m2_pupil)

@@

## Final Pupil Dilation Model Plot

>>>>

library(rio) # install.packages("rio")
library(tidyverse) # install.packages("tidyverse")
library(mgcv) # install.packages("mgcv")
library(itsadug) # install.packages("itsadug")
library(irr) # install.packages("irr")
library(ggpubr) # install.packages("ggpubr")
library(gt) # install.packages("gt")
library(gtsummary) # install.packages("gtsummary")
library(glmmTMB) # install.packages("glmmTMB")
library(performance) # install.packages("performance")
library(sjPlot) # install.packages("sjPlot")

@@

Adding in AR(1) structure to address autocorrelation concerns. Bumped up knots to 25.

@@

summary.gam(m1_pupil)

m1_prediction <- plot_smooth(m1_pupil,
            view = "time_norm",
            plot_all = "condition",
            rug = F, se = 1, rm.ranef = T)

@@

In this model, the scaled t distribution was added to the model specifications to fix model fit to the data. The model was originally running into an error when the random smooth for phrase was included in the specification. Therefore, it was removed in this model. The random smooth for listener remained in this model.

m2_pupil <- bam(normed_pupil ~ condition +
                  s(time_norm, by = condition, k = 25) +
                  # Random Smooth for listener
                  s(time_norm, listener, bs = 'fs', m = 1),
                discrete = T,
                AR.start = start_event,
                rho = start_value_rho(m0_pupil),
                family = "scat",
                data = pupil_df)

summary.gam(m2_pupil)

@@

### Model 3

In model 3, we are testing whether adjusting the random smooth specification significantly improves model fit to the data. In model 2, there was just a random smooth for listener. This random smooth specification is updated here to be a random smooth per listener for each condition level (interactions between speaker and phrase repetition accuracy).

m3_pupil <- bam(normed_pupil ~ condition +
                  s(time_norm, by = condition, k = 25) +
                  # Random Smooths for listener
                  s(time_norm, listener, by = condition, bs = 'fs', m =1),
                discrete = T,
                AR.start = start_event,
                rho = start_value_rho(m0_pupil),
                family = "scat",
                data = pupil_df)

summary.gam(m3_pupil)

m3_prediction <- plot_smooth(m3_pupil,
            view = "time_norm",
            plot_all = "condition",
            rug = F, se = 1, rm.ranef = T)

The model residuals look similar. The adjusted R-squared value is .192.

gam.check(m3_pupil)

acf_resid(m3_pupil)

## Model Comparison

To test whether model 3 significantly improved the fit to the data, a generalized likelihood ratio test was conducted. The results suggest that model 3 did not significantly improve model fit. Therefore, we will retain model 2.

compareML(m2_pupil, m3_pupil, signif.stars = T, suggest.report = T)

## Final Model (Model 2)
